Exploring probability distributions for bivariate temporal granularities

### 2019-12-09 main.Rmd
minor changes till cyclic granularities fixed

The cyclic calendar algebra consists of two kinds of operations:  
(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities

@@

\begin{equation} \label{eq:eq7}
\begin{split}
C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z)-1)) \\  
  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z) - 1) - 1]) \\
  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}} - 1))\\
  & =  L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z) - 1))\\
  &\vdots\\
  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1))\\
\end{split}
\end{equation}

*Example:*Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar..

@@

Let us revisit Gregorian calendar for addressing this case. Suppose we have a hierarchy table using some linear granularties from Gregorian calendar. Since months consists of unequal number of days, any linear granularity unit with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities. The single-order-up granularity day_month is quasi-circular. Any single-order-up granularities which are formed by linear granularities below days are circular. Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.

@@

|  G	        |C                 |P
|------------	|------------------|----------------|
|  minute     |minute-of-hour    |  60            |
|  hour    	  |hour-of-day       |  24  	        |
|  day      	|day-of-month      |"quasi-circular"|
|  month    	|month-of-year     |  12            |
|  year   	  | 1                |	1             |

@@

where, $T_w$ are the sets of indices of linear granularity day such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.

@@

\begin{equation} \label{eq:multifromsingle-quasi}
\begin{split}
C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)) \\ 
& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z) - 1)) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)\\
\end{split}
\end{equation}

@@

A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**. And the ones that do not lead to any missing combinations are referrednto as **harmonies**.

@@

Estimating the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation. The measured variable in the considered data structure can be dicrete or continuous. The probability distribution of them could be studied by estimating the probability density function (continuous) or probability mass function (dicrete). The approach can be parametric or non-parametric. A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself. These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses. The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles. Severalforms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:

>>>>

The cyclic calendar algebra consists of broadly two kinds of operations:  
(1) **single-to-multiple:** representation of multiple-order-up cyclic granularities from single order-up cyclic granularities  
(2) **multiple-to-single:** representation of single-order-up cyclic granularities from multiple order-up cyclic granularities

@@

\begin{equation} \label{eq:eq7}
\begin{split}
C_{(G_l,G_m)}(z) & = L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{(G_(l+1),G_m)}(z))) \\  
  & =   L(C_{(G_l,G_{l+1}}(z) + k(l, l+1)[ C_{G_{l+1},G_{l+2}}(z) + k(l+1, l+2)( C_{G_{l+2}, G_m}(z))) \\
  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+1)k(l+1, l+2)(C_{G_{l+2}, G_{m}}))\\
  & =  L(C_{G_l,G_{l+1}}(z) + k(l, l+1)(C_{G_{l+1},G_{l+2}}(z) - 1) + k(l, l+2)(C_{G_{l+2},G_{l+m}}(z)))\\
  &\vdots\\
  & = L(\sum_{i=0}^{m - l - 1} k(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z)))\\
\end{split}
\end{equation}

*Example:* Let us use the equation \ref{eq:eq7} to compute the multiple-order-up granularity uinal_katun for Mayan calendar.

@@

Let us revisit Gregorian calendar for addressing this case. Suppose we have a hierarchy table using some linear granularties from Gregorian calendar. Since months consists of unequal number of days, any linear granularity with higher order than months will also have unequal number of days. This is an example of a hierarchy structure which has both circular and quasi-circular single-order-up granularities. The single-order-up granularity day-of-month is quasi-circular. Any single-order-up granularities which are formed by linear granularities below days are circular. Similarly, all single-order-up granularities which are formed using linear granularities with orders higher than months are also circular.

@@

|  G	        |C                 |P
|------------	|------------------|----------------|
|  minute     |minute-of-hour    |  60            |
|  hour    	  |hour-of-day       |  24  	        |
|  day      	|day-of-month      |quasi-circular  |
|  month    	|month-of-year     |  12            |
|  year   	  | 1                |	1             |

@@

where, $T_w$ is such that $month(w) = \bigcup_{z \in T_w}day(z)$ and $\vert T_w \vert$ is the cardinality of set $T_w$.

@@

\begin{equation} \label{eq:multifromsingle-quasi}
\begin{split}
C_{G_l,G_m}(z) & = L(C_{G_{l},G_{m'}}(z) + P(l, m')(C_{G_{m'},G_{m}}(z)-1)) \\ 
& = L(\sum_{i=0}^{m' - l - 1} P(l, l+i)(C_{G_{l+i},G_{l+i+1}}(z))) + P(l, m')(C_{G_{m'},G_{m}}(z))\\
\end{split}
\end{equation}

@@

A pair of cyclic granularities which lead to structurally, event-driven or build-based empty-combinations are referred to as to as **clashes**. And the ones that do not lead to any missing combinations are referred as **harmonies**.

@@

Estimating the distribution of the measured variable can give valuable indication of features like central tendency, skewness, multimodality, tail behavior or variation. The measured variable in the considered data structure can be dicrete or continuous. The probability distribution of them could be studied by estimating the probability density function (continuous) or probability mass function (dicrete). The approach can be parametric or non-parametric. A non-parametric approach makes less rigid assumptions about the distribution of the observed data making the data speak more for itself. These resonate more with the notion of EDA which advocates exploring data for patterns and relationships without requiring prior hypotheses. The entire distribution could be estimated using different density estimation methods or summarized using empirical quantiles. Several forms of  non-parametric density estimators and quantile estimators exist in the literature, the most commonly used form ( @Silverman1986-cl,  @Jones1992-pj) are as follows:
